Spikes demo

### 2018-08-23 analysis/spikesdemo.Rmd
wflow_publish("spikesdemo.Rmd")

This script illustrates key features of the smash method on a small,
simulated data set.

@@

```{r load-pkgs}
library(ashr)
library(smashr)
library(scales)
library(ggplot2)
suppressMessages(library(cowplot))
source("../code/misc.R")
```

@@

Function `spike.f` defines the mean signal.

@@

*Add text here.*

@@

```{r plot-simulated-data}
par(cex.axis = 0.8,cex.sub = 1,cex.lab = 1)
plot(mu.sp,type = 'l',ylim = c(-0.05,1),xlab = "position",
     ylab = "",lwd = 1.7,xaxp = c(0,1024,4),yaxp = c(0,1,4))
lines(mu.sp + 2*sig.cb,col = "darkorange",lty = 5,lwd = 1.8)
lines(mu.sp - 2*sig.cb,col = "darkorange",lty = 5,lwd = 1.8)
points(x.sim,cex = 0.7,pch = 16,col = "darkblue")
```

@@

Apply smash and translation invariant (TI) thresholding to the
"spikes" data. Here, we run the TI thresholding twice---once when
the standard deviation (s.d.) function is provided, and once when it
is estimated using the MAD algorithm.

```{r denoise-signals}
sig.est  <- sqrt(2/(3 * (n - 2)) *
                 sum((1/2 * x.sim[1:(n-2)] - x.sim[2:(n-1)] + x.sim[3:n])^2/2))
mu.smash <- smash(x.sim,family = "DaubLeAsymm",filter.number = 8)
mu.ti    <- ti.thresh(x.sim,method = "rmad",family = "DaubLeAsymm",
                      filter.number = 8)
mu.ti.homo <- ti.thresh(x.sim,sigma = sig.est,family = "DaubLeAsymm",
                        filter.number = 8)
```

Get the wavelet coefficients and their variances.

@@

Plot the distribution of observed wavelet coefficients.

```{r}
par(cex.axis = 0.8,cex.lab = 0.8)
hist(wc.sim[4,],breaks = 2,xlab = "observed wavelet coefficients",
     xlim = c(-25,25),ylim = c(0,600),col = "darkblue",xaxp = c(-25,25,10),
     yaxp = c(0,600,6),main = "")
hist(wc.sim[10,],breaks = 40,add = TRUE,col = "darkorange")
```

@@

```{r}
par(cex.axis = 0.8,cex.lab = 0.8)
plot(c(),c(),xlab = "observed wavelet coefficients",
     ylab = "shrunken wavelet coefficients",
     xlim = c(-2.5,2.5),ylim = c(-2.5,2.5))
abline(0,1,lty = 1,col = "gray",lwd = 1)
points(wc.sim[10,],wc.sim.shrunk[[9]]$PosteriorMean,pch = 20,cex = 0.6,
       col = "darkorange")
points(wc.sim[4,],wc.sim.shrunk[[3]]$PosteriorMean,pch = 20,cex = 0.6,
       col = "darkblue")
```

@@

```{r}
wc.sig.3 <- 1/wc.pres[[3]]
p <- ggplot(data.frame(observed = wc.sim[4,],
                       shrunken = wc.sim.shrunk[[3]]$PosteriorMean,
                       se       = wc.sig.3),
            aes(x = observed,y = shrunken,col = se)) +
  geom_point() +
  xlim(c(-1,1)) +
  ylim(c(-1,1)) +
  scale_color_gradientn(colors = c("deepskyblue","darkblue")) +
  theme_cowplot()
print(p)
```

>>>>

This script illustrates key features of the SMASH method on a small,
simulated data set with heteroskedastic variances.

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,fig.align = "center")
```

@@

```{r load-pkgs, warning=FALSE}
library(ashr)
library(smashr)
library(scales)
library(ggplot2)
suppressMessages(library(cowplot))
source("../code/misc.R")
```

@@

Variable `mu.sp` defines the mean signal.

@@

Simulate the noisy observations from the normal distribution with
means that are given by the "spikes" function, and with
spatially-structured variances.

@@

```{r plot-simulated-data, fig.height=5, fig.width=10}
par(cex.axis = 1,cex.lab = 1.5)
plot(mu.sp,type = 'l',ylim = c(-0.05,1),xlab = "position",
     ylab = "",lwd = 1.7,xaxp = c(0,1024,4),yaxp = c(0,1,4))
lines(mu.sp + 2*sig.cb,col = "darkorange",lty = 5,lwd = 1.8)
lines(mu.sp - 2*sig.cb,col = "darkorange",lty = 5,lwd = 1.8)
points(x.sim,cex = 0.7,pch = 16,col = "darkblue")
```

This plot shows the "spikes" mean function (the black line),
+/- 2 standard deviations (the yellow lines), and
the simulated data (the blue points).

@@

Now we apply SMASH, as well as the translation invariant (TI)
thresholding method, to the "spikes" data. Here we run the TI
thresholding method twice---once when the variance is assumes to be
constant ("homoskedastic"), and once when it is estimated using the
RMAD algorithm. The first method yields very poor estimates, so they
are not shown in the plots below.

The SMASH method is able to estimate the variance.

```{r denoise-signals}
sig.est  <- sqrt(2/(3 * (n - 2)) *
              sum((1/2 * x.sim[1:(n-2)] - x.sim[2:(n-1)] + x.sim[3:n])^2/2))
mu.smash <- smash(x.sim,family = "DaubLeAsymm",filter.number = 8)
mu.ti.homo <- ti.thresh(x.sim,sigma = sig.est,family = "DaubLeAsymm",
                        filter.number = 8)
mu.ti    <- ti.thresh(x.sim,method = "rmad",family = "DaubLeAsymm",
                      filter.number = 8)
```

It should take at the very most a few minutes to run SMASH and TI
thresholding.

Get the (true) wavelet coefficients and their variances.

@@

Plot the distribution of the observed wavelet coefficients at a
coarser scale (scale = 1, orange) and at a finer scale (scale = 7,
dark blue).

```{r plot-wc-histogram, fig.height=4, fig.width=6}
par(cex.axis = 1,cex.lab = 1.5)
hist(wc.sim[4,],breaks = 2,xlab = "observed wavelet coefficients",
     xlim = c(-25,25),ylim = c(0,600),col = "darkblue",xaxp = c(-25,25,10),
     yaxp = c(0,600,6),main = "")
hist(wc.sim[10,],breaks = 40,add = TRUE,col = "darkorange")
```

Observe that the scale = 7 WCs are much more concentrated near zero
because the signal is smoother at this finer scale.

@@

```{r wc-scatterplot, fig.height=5, fig.width=5}
par(cex.axis = 1,cex.lab = 1.5)
plot(c(),c(),xlab = "observed wavelet coefficients",
     ylab = "shrunken wavelet coefficients",
     xlim = c(-2.5,2.5),ylim = c(-2.5,2.5))
abline(0,1,lty = 1,col = "gray",lwd = 1)
points(wc.sim[10,],wc.sim.shrunk[[9]]$PosteriorMean,pch = 20,cex = 1,
       col = "darkorange")
points(wc.sim[4,],wc.sim.shrunk[[3]]$PosteriorMean,pch = 20,cex = 1,
       col = "darkblue")
```

@@

```{r}
wc.sig.3 <- 1/wc.pres[[3]]
p <- ggplot(data.frame(observed = wc.sim[4,],
                       shrunken = wc.sim.shrunk[[3]]$PosteriorMean,
                       se       = wc.sig.3),
            aes(x = observed,y = shrunken,col = se)) +
  geom_point(na.rm = TRUE) +
  xlim(c(-1,1)) +
  ylim(c(-1,1)) +
  scale_color_gradientn(colors = c("deepskyblue","darkblue")) +
  theme_cowplot()
print(p)
```

The scale = 7 WCs are strongly shrunk toward zero, whereas the scale =
1 WCs are not shrunk nearly as much; SMASH infers from the data that
the scale = 7 WCs are heavily concentrated around zero, and
consequently SMASH shrinks them more strongly.
